### 2010-02-15 doc/slides.tex
*** empty log message ***

\documentclass{beamer}
%\documentclass[handout]{beamer}

%\usepackage{beamerjcf}
\usepackage[latin1]{inputenc}

\definecolor{kwblue}{rgb}{0.67,0.12,0.92}
\definecolor{ceruleanblue}{rgb}{0, 0.48, 0.65}
\definecolor{lightpink}{rgb}{1., 0.71, 0.75}
\definecolor{lightblue}{rgb}{0.8,0.8,1}
\definecolor{lightred}{rgb}{1,0.8,0.8}

\let\emph\alert

\begin{document}

\title{Map/Reduce for the common man}
\author[Jean-Christophe]{Jean-Christophe Filli\^atre \& Kalyan Krishnamani}
\institute{
}
\date{ProVal, February 19, 2010}

\begin{frame}
  \titlepage
  \pgfimage[height=8mm]{cnrs-logo2}\hfill
  \pgfimage[height=6mm]{saclay}\hfill
  \pgfimage[height=10mm]{lrilogo}\hfill
  \pgfimage[height=8mm]{upsudlogo}
\end{frame}

\begin{frame}\frametitle{Overview}
  \begin{enumerate}
  \item Google's Map/Reduce
  \item Ocaml's Map/Reduce
  \item Implementation details
  \end{enumerate}
\end{frame}

\begin{frame}\frametitle{Google's Map/Reduce}
  what it is

  what it offers

  application domains

\end{frame}

\begin{frame}
  \begin{center}
    \hrulefill\\
    \emph{Ocaml's Map/Reduce}
  \end{center}
\end{frame}

\begin{frame}\frametitle{}
  two ideas
  \begin{itemize}
  \item polymorphic map/fold instead of map/reduce
  \item sequential / cores / network
  \end{itemize}
\end{frame}

\begin{frame}\frametitle{}
  brief overview of the API
\end{frame}

\begin{frame}\frametitle{}
  first demo
\end{frame}

\begin{frame}\frametitle{}
  detailed API
\end{frame}

\begin{frame}\frametitle{}
  another, more impressive demo ?
\end{frame}

\begin{frame}
  \begin{center}
    \hrulefill\\
    \emph{Implementation details}
  \end{center}
\end{frame}

\begin{frame}\frametitle{}
  common part between cores and network

  output/input\_value on core

  architecture-independent protocol for network

  same program for master and workers
\end{frame}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: