### 2010-05-25 doc/icfp-2010/response.txt
*** empty log message ***

Response to review #5A:
-----------------------

> The major weakness however is that the task creation is
> completely static, i.e., all tasks must be created ahead of time.

It is true that in the derived API (Figure 1 on page 4) the tasks are
completely static. But in the low-level API, new tasks can be
generated dynamically, as mentioned on page 2, first paragraph (right
below the type of the master function).

> The author claims that strategy utilizes efficient storage of data

Actually, we did not, and only mentioned that as an interesting
extension to our library (last paragraph of Section 7, first bullet).

Response to review #5B:
-----------------------

> I wondered why oc4mc wasn't used for the single machine version

As far as we understand, oc4mc is a modification of OCaml (available
for a single architecture). On the contrary, our library does not
require any modification of the compiler.

Response to review #5C:
-----------------------

> (also, the example code is heavily imperative with lots of unnecessary refs).

It is true that the example code in Section 3 is making use of
imperative features. The reason is that it uses the low-level API of
Section 2. However, the derived API of Section 4 provides fold-like
functions and thus allows pure functional code, even if we did not show
in the paper.

> the connections to functional programming are tenuous

Our library API is mostly composed of higher-order functions: in the
low-level API (Section 2), master is a higher-order function; in the
derived API (Figure 1), all functions are higher-order map/fold.

Additionally, our library is heavily relying on OCaml module
system: first, to provide three different implementations with the
same interface; second, to achieve maximal code reuse in the library
itself using functors (as explained at the end of page 3 for instance).

Responses for explicit questions for authors:
---------------------------------------------

> What happens if a worker raises an exception?

It is considered as a worker failure. We leave it to the user to catch
it and handle it (using a sum type for instance).

> It appears that your local implementation does not handle faults in the
> workers (e.g., if you run out of a system resource). Is that the case?

Yes, it is. One possibility to overcome this would be to change the
resource limits in the user worker function (right after the fork).

> In the network implementation, how to you initiate the workers?

As of now, we do not provide any facility in order to do this.
Currently, we assume the user deploys the workers manually.