### 2010-07-01 doc/ml-workshop-2010/main.tex
*** empty log message ***

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

\documentclass[preprint]{sigplanconf}

% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath,color,url}

\newcommand{\Ocaml}{OCaml}
\newcommand{\functory}{\textsf{Functory}}
\newcommand{\JoCaml}{Jo{\&\!}Caml}
\newcommand{\unix}{\textsc{Unix}}

\begin{document}

\conferenceinfo{Workshop on ML 2010}{September 26, 2010, Baltimore, Maryland.} 
\copyrightyear{2010} 
\copyrightdata{[to be supplied]}

\titlebanner{submitted to Workshop on ML 2010}        % These are ignored unless
\preprintfooter{Distributed Computing for the Common Man}   % 'preprint' option specified.

\title{Distributed Computing for the Common Man}
%\subtitle{Subtitle Text, if any}

\authorinfo{Jean-Christophe Filli\^{a}tre}
           {CNRS, LRI, Univ Paris-Sud 11, Orsay F-91405\\
            INRIA Saclay - \^{I}le-de-France, ProVal, Orsay, F-91893}
           {filliatr@lri.fr}
\authorinfo{K. Kalyanasundaram}
           {ProVal, INRIA Saclay - \^{I}le-de-France, Orsay, F-91893}
           {kalyan.krishnamani@inria.fr}

\maketitle

\begin{abstract}
  We present a distributed computing library for
  functional programming languages. The main features of this library
  include (1) a generic API, (2) several implementations to
  adapt to different computing environments such as sequential,
  multi-core or network, and (3) a reliable fault-tolerance mechanism.
  This paper describes the motivation behind this work, as well as
  the design and implementation of the library. It also demonstrates
  the potential of the library using realistic experiments.
\end{abstract}

% \category{D1.1}{Programming techniques}{Applicative (Functional) Programming}

% \terms
% Distributed Computing, Functional Programming

%\keywords

\section{Introduction}

This paper introduces a generic library for distributed computing. This
work was initially motivated by the computing needs that exist in our
own research team. Our applications include large-scale deductive
program verification, which amounts to checking the validity of a
large number of logical formulas using a variety of automated theorem
provers~\cite{filliatre07cav}. Our computing infrastructure consists
of a few powerful multi-core machines (typically 8 to 16 cores) and
several desktop PCs (typically dual-core). However, for our
application needs, there is no library that helps in exploiting such a
computing infrastructure in our favorite functional programming
language.  Hence we designed and implemented such a
library\footnote{This library is implemented in \Ocaml, but the
  implementation should be straightforward in any functional
  programming language.}, which is the subject of this paper.

The distributed computing library presented in this paper is not a
library that helps in parallelizing computations. Rather, it provides
facilities for reliable, distributed execution of parallelizable
computations. In particular, it provides a set of APIs that allows the
execution of large-scale parallelizable computations, very relevant to
our application needs (and presumably relevant to a variety of
real-world applications), over multiple cores in the same machine or
over a network of machines. 
The most important features of our library are the following:
\begin{itemize}
\item \emph{Genericity}: 
  it allows various patterns of polymorphic computations;
\item \emph{Simplicity}: switching between multiple cores on the same
  machine and a network of machines is as simple as changing a couple
  of lines of code;
\item \emph{Task distribution and fault-tolerance}: 
  it relieves the user from implementing task distribution routines
  and reliable fault-tolerance mechanisms.
\end{itemize}
It is worth noting that the library is not targeted at applications
running on server farms, crunching enormous amounts of data. Neither
is it limited to research labs possessing few powerful
machines. It serves a variety of users and a wide
spectrum of needs, from desktop PCs to networks of machines, and hence
the title.

In addition to the above, distributed computing environments also
implement mechanisms for fault tolerance, efficient storage, and
distribution of tasks. This is required to handle network failures
that may occur, as well as to optimize the usage of machines in the
network. Another concern of importance is the transmission of messages
over the network. This requires efficient 
\emph{marshaling} of data, that is encoding and decoding of data 
for transmission over different computing environments.  It is desirable to
maintain architecture independence while transmitting marshalled data,
as machines in a distributed computing environment often run on
different hardware architectures and make use of varying software
platforms. For example, machine word size or endianness may be different
across machines on the network.

\paragraph{A Functional Programming Approach.}
Our work was initially inspired by Google's
MapReduce\footnote{Ironically, Google's approach itself was inspired
  by functional programming primitives.}. However, our functional
programming environment allows us to be more generic. 
The main idea behind our approach is that
workers may implement any polymorphic function:
\begin{ocaml}
  worker: 'a -> 'b
\end{ocaml}
where \of{'a} denotes the type of tasks and \of{'b} the type of results.
Then the master is given a list of initial tasks, together with a
function to handle the results:
\begin{ocaml}
  master: ('a -> 'b -> 'a list) -> 'a list -> unit
\end{ocaml}
The function passed to the master is applied whenever a result is
available and may in turn generate new tasks (hence the return type
\of{'a list}).  The master is executed as long as there are pending
tasks.

Our library makes use of \Ocaml's marshaling capabilities as much as
possible. Whenever master and worker executables are exactly the same,
we can marshal polymorphic values and closures. However, it is not
always possible to have master and workers running the same
executable. In this case, we cannot marshal closures anymore but we
can still marshal polymorphic values as long as the same version of
\Ocaml\ is used to compile master and workers. When different versions
of \Ocaml\ are used, we can no longer marshal values but we can still
transmit strings between master and workers. Our library adapts to all
these situations, by providing several APIs.

\medskip
The paper is organized as follows. In Section~\ref{sec:API}, we
describe the main idea behind the generic API, and introduce
\of{master} and \of{worker} programs. Section~\ref{sec:example}
illustrates the API usage with an example.
In Section~\ref{sec:derived}, we
derive some interesting functions from the generic
API. Section~\ref{sec:implem} delves into the implementation details
of our library, while Section~\ref{sec:experiments} illustrates the
potential of the presented library through experimental evaluation. We
compare our approach with relevant related work in
Section~\ref{sec:future} and outline our future work.

Our library, \functory, is implemented in \Ocaml\ and is available
from \url{http://www.lri.fr/~filliatr/functory/}.

TODO : refer to the technical report available on-line

\section{API}\label{sec:API}

Our work was initially inspired by Google's
MapReduce\footnote{Ironically, Google's approach itself was inspired
  by functional programming primitives.}, from where we borrow some 
terminology:
\begin{itemize}
\item A notion of \emph{tasks} which denote atomic computations
  to be performed in a distributed manner; 
\item A set of processes (possibly executing on remote machines)
  called \emph{workers} that perform
  the tasks, producing results;
\item A single process called a \emph{master} which is in charge
  of distributing the tasks among the workers and managing results
  produced by the workers.
\end{itemize}
The main function of our API combines these three notions of tasks,
master and workers into a single higher-order polymorphic function
\of{compute}:
\begin{ocaml}
  val compute : 
    worker:('a -> 'b) -> 
    master:('a * 'c -> 'b -> ('a * 'c) list) -> 
    ('a * 'c) list -> unit
\end{ocaml}
Tasks are pairs, of type \of{'a * 'c}, where the first component is
passed to the worker and the second component is local to the master.
The \ocaml{worker} function should be pure and is executed in parallel
in all worker processes. The function \ocaml{master}, on the
contrary, can be impure and is only executed in the master process.
The \ocaml{master} function typically stores results 
in some internal data structure.

Our library provides not just a single \ocaml{compute}
function as above, but instead five different versions
depending on the execution context:
\begin{enumerate}
\item \textbf{Purely sequential execution:}
  this is mostly intended for debugging;

\item \textbf{Several cores on the same machine:} 
  this implementation makes use of \unix\ processes and 
  provides a polymorphic \ocaml{compute} function;

\item \textbf{Same executable run on master and worker machines:}
  this implementation makes use of the ability to marshal \Ocaml\
  closures and  provides a polymorphic \ocaml{compute}
  function.
  Depending on whether the program is run as a master or as a worker,
  the relevant arguments of \ocaml{compute} are used.

\item \textbf{Master and workers are different programs, compiled with
    the same version of \Ocaml:} 
  we can no longer marshal closures but we can still
  marshal polymorphic values. As a consequence, 
  the \ocaml{compute} function is split into two
  polymorphic functions, to implement the master and workers 
  respectively:%
\vspace{-1em}
\begin{ocaml}
val Worker.compute : ('a -> 'b) -> unit -> unit
val Master.compute : 
  ('a * 'c -> 'b -> ('a * 'c) list) -> 
  ('a * 'c) list -> unit
\end{ocaml}
\item \textbf{Master and workers are different programs, not even
    compiled with the same version of \Ocaml:} we can no
  longer use marshaling, so the
  \ocaml{compute} function is split into two monomorphic functions
  over strings:%
\vspace{-1em}
\begin{ocaml}
val Worker.compute : (string -> string) -> unit -> unit
val Master.compute : 
  (string * 'c -> string -> (string * 'c) list) -> 
  (string * 'c) list -> unit
\end{ocaml}
  Note that the second component of each task is still polymorphic (of
  type \ocaml{'c} here), since it is local to the master.
\end{enumerate}

% Our library is organized into three modules: \of{Sequential} for the
% pure sequential implementation, \of{Cores} for multiple cores on the
% same machine and \of{Network} for a network of machines, respectively.
% The \of{Network} module itself is organized into three sub-modules,
% called \of{Same}, \of{Poly} and \of{Mono},
% corresponding to contexts 3, 4 and 5 above. 
% The next section demonstrates the use of these modules.

  \begin{ocaml}
    val map : 
      f:('a -> 'b) -> 'a list -> 'b list 
    val map_local_fold : 
      map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
      'c -> 'a list -> 'c 
    val map_remote_fold : 
      map:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 
      'c -> 'a list -> 'c 
    val map_fold_ac : 
      map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
      'b -> 'a list -> 'b 
    val map_fold_a : 
      map:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 
      'b -> 'a list -> 'b
  \end{ocaml}

\section{Illustrative Example}\label{sec:example}

Let us assume we need to compute a sum such as
\begin{displaymath}
  s(n) = \sum_{i=0}^{n}f(i)
\end{displaymath}
where $f$ is a pure and computationally expensive function.
We start with the sequential implementation provided by the library:
\begin{ocaml}
  open Sequential
\end{ocaml}
The idea is to test our code on small values of $n$ before
parallelizing the computation for larger values.
Let us assume that $f$ is implemented as a top-level \of{worker} function:
\begin{ocaml}
  let worker i = ...
\end{ocaml}
and that $n$ is obtained from the command line:
\begin{ocaml}
  let n = int_of_string Sys.argv.(1)
\end{ocaml}
The list of tasks is simply the list of integers $0,1,\dots,n$,
computed as follows:
\begin{ocaml}
  let tasks = 
    let l = ref [] in 
    for i = 0 to n do l := (i, ()) :: !l done; 
    !l
\end{ocaml}
Each task is actually a pair, where the second component contains
some information which is kept local to the master process.
It is irrelevant here and thus the value \of{()} of type \of{unit} is used.
The current sum is kept into a global reference \of{s}:
\begin{ocaml}
  let s = ref 0
\end{ocaml}
The master function receives a task \of{(i,())} together with its
result \of{fi} and simply adds \of{fi} to \of{s}. There is no new task
generated, hence the returned empty list.
\begin{ocaml}
  let master (i,()) fi = s := !s + fi; []
\end{ocaml}
Finally, the whole computation takes place when we call the
\of{compute} function:
\begin{ocaml}
  let () = compute ~worker ~master tasks
\end{ocaml}

Now, let us assume we want to make use of a 4-core machine to compute
$s(n)$. It is as easy as just replacing the very first line with
\begin{ocaml}
  open Cores
  let () = set_number_of_cores 4
\end{ocaml}
while the rest of the code remains unchanged.

Later, we may want to use a network of machines instead, for example
two machines called \texttt{orcus} and \texttt{belzebuth} with 4 and 8
cores respectively. Similar to the above, we just replace the first
lines with
\begin{ocaml}
  open Network
  let () = declare_workers ~n:4 "orcus"
  let () = declare_workers ~n:8 "belzebuth"
  open Same
\end{ocaml}
Here, \of{Same} is a module which is to be used when master and worker
are running the same executable. It provides a \of{compute} function
which has the same signature as in modules \of{Sequential} and
\of{Cores}, so that the rest of the code remains unchanged. Master and
worker processes are distinguished at run-time using an environment
variable \texttt{WORKER} which is set/unset.

If we need to write two different programs for master and worker, for
reason of binary incompatibility or any other reason, the library API
is providing functions to do that. If master and worker are still
compiled with the same version of \Ocaml, we use the \of{Poly} module
which provides a polymorphic API. Let us start with the worker program.
It is as simple as
\begin{ocaml}
  open Poly
  let worker i = ...
  let () = Worker.compute worker ()
\end{ocaml}
The \of{Worker.compute} function enters a loop which waits for
tasks sent by the master and returns results computed using \of{worker}.
The master program is almost the same as before. First, we replace
module \of{Same} with module \of{Poly}:
\begin{ocaml}
  open Network
  let () = declare_workers ~n:4 "orcus"
  let () = declare_workers ~n:8 "belzebuth"
  open Poly
\end{ocaml}
Tasks and \of{master} function are unchanged:
\begin{ocaml}
  let tasks = ...
  let s = ref 0
  let master (i,()) fi = ...
\end{ocaml}
Finally, we start the computation with \of{Master.compute}, which does
not have a \of{worker} parameter anymore:
\begin{ocaml}
  let () = Master.compute ~master tasks
\end{ocaml}

When master and worker programs are compiled with different
versions of the \Ocaml\ compiler, our library still provides
a monomorphic API over strings. As a consequence, we need to convert
integers to and from strings in both master and worker.
The modified worker program is as follows:
\begin{ocaml}
  open Mono
  let worker i = ...
  let worker_string i = string_of_int (worker (int_of_string i))
  let () = Worker.compute worker_string ()
\end{ocaml}
The master program is modified in a similar way.
We simply replace \of{Poly} with \of{Mono} and encode/decode integers
as strings, as follows:
\begin{ocaml}
  let tasks = 
    let l = ref [] in 
    for i = 0 to n do l := (string_of_int i, ()) :: !l done; 
    !l
  let master (i,()) fi = s := !s + int_of_string fi; []
\end{ocaml}

\section{Experiments}\label{sec:experiments}

Here we demonstrate the potential of our library for using the
classical $N$-queens problem and large-scale verification using SMT-solvers.

\paragraph{N-queens.} 
The following table shows execution times\footnote{machine
  characteristics: 8 cores Intel Xeon 3.2 GHz running Debian Linux.} 
in seconds for various values of $N$
and our three different implementations (sequential, 8 cores on a
single machine and a network of 8 workers).
\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    N & \#tasks  & sequential& cores                 & network 
    \\\hline\hline
    16 &   16    &  15.2     &   2.04 (7.45$\times$) &  2.35  (6.47$\times$) 
    \\\hline
       &  210    &  15.2     &   2.01 (7.56$\times$) & 21.80  (0.69$\times$)
    \\\hline
    17 &   17    & 107.0     &  17.20 (6.22$\times$) & 16.20  (6.60$\times$)
    \\\hline
       &  240    & 107.0     &  14.00 (7.64$\times$) & 24.90  (4.30$\times$)
    \\\hline
    18 &   18    & 787.0     & 123.00 (6.40$\times$) & 125.00 (6.30$\times$)  
    \\\hline
       &  272    & 787.0     & 103.00 (7.64$\times$) & 124.00 (6.34$\times$)  
    \\\hline
    19 &   19    &6120.0     & 937.00 (6.53$\times$) & 940.00 (6.51$\times$)  
    \\\hline
       &  306    &6130.0     & 796.00 (7.70$\times$) & 819.00 (7.48$\times$)
    \\\hline
  \end{tabular}
\end{center}
% From the table above, it is clear that the \of{Cores} and \of{Network}
% implementations provide a significant speedup. As evident from the
% last row, the speedup is almost 8, which is also the number of
% cores we use.  It is also evident from the last column that the
% \of{Network} implementation performs significantly better when the
% computation time dominates in the total execution time.  The two extreme
% cases correspond to the second and the last row: in the second row, the
% communication time dominates and is in fact more than 91\%\ of the
% total execution time; on the other hand, for the last row
% communication time amounts to just 4.6\%\ of the total execution time.
% As expected, the network implementation is only beneficial when the
% computation time for each individual task is significant, which is the
% case in realistic examples.

\paragraph{SMT Solvers.}
Our setup consists of 3 machines with 4, 8 and 8 cores
respectively and the benchmark consists of
80 challenging verification conditions (VC).  
The purpose of the experiment is to check the validity of each VC
using the four SMT-solvers Alt-Ergo, Simplify, Z3 and CVC3.
With a timeout of 1 minute, each task completes 
with one of the four possible outcomes: \emph{valid},
\emph{unknown} (depending on
whether the VC is valid or undecided by the prover), 
\emph{timeout} and \emph{failure}.
Timings in minutes are the following:
\begin{center}
  \begin{tabular}{|r||r|r|r|r|}
    \hline
    prover   & valid & unknown & timeout & failure
    \\\hline\hline
    Alt-ergo & 406.0 & 3.0   &  11400.0 & 0.0       
    \\\hline
    Simplify &  0.5   & 0.4   &  1200.0 & 222.0   
    \\\hline
    Z3       & 80.7   & 0.0   &  1800.0 & 1695.0   
    \\\hline
    CVC3     & 303.0  & 82.7  &  4200.0 & 659.0   
    \\\hline
  \end{tabular}
\end{center}
These figures sum up to more than 6 hours if provers were executed
sequentially. However, using our library and our 3-machine
infrastructure, it completes in 22 minutes and 37 seconds, giving us a
speedup of more than 16$\times$.

\end{document}

% LocalWords:  parallelize functor parameterized indices endianness monomorphic
% LocalWords:  genericity executables OCamlMPI MapReduce